---
title: "Bootstrapping"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(modelr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```


## Simulate data
```{r}
n_samp = 250

sim_df_const =
  tibble(
    x = rnorm(n_samp, 1, 1),
    error = rnorm(n_samp, 0, 1),
    y = 2 + 3 * x + error
  )

sim_df_nonconst =
  sim_df_const %>% 
  mutate(
    error = error * .75 * x,
    y = 2 + 3 * x + error
  )
```

Plot the datasets
```{r}
sim_df_const %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm")
#Equal variances. Underlying assumption for linear regression

sim_df_nonconst %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = "lm")
#Unequal variance.  When x close to zero, residuals are very very small.  When x is far from zero, the residuals spread out.  Can still do linear regression, but one of the key assumptions - equal variance - is not met.  So difficult to make inferences - specifically difficult to construct confidence intervals.
#We will try to solve this problem using bootstrapping
```

```{r}
lm(y ~ x, data = sim_df_const) %>%  broom::tidy()
lm(y ~ x, data = sim_df_nonconst) %>%  broom::tidy()
```

Bootstrap
Try to solve this issue of unequal variances with the bootstrap
What if I had another dataset like this one - get intercept and slope
Repeat
Repeat
Could I just then look and see what is the actual variance of the intercept and slope under repeated sampling
This is what we are trying to do with the bootstrap

## Draw one bootstrap sample
First, write a bootstrap function
The reason for this is so that later, can repeat the function

```{r}
boot_sample = function(df) {
  
  sample_frac(df, replace = TRUE) %>% 
#Why bootstrap sample should be the same size as original sample?  Because confidence intervals (variance estimates) are very sensitive to sample size.  Fewer point estimate = wider confidence interval and vice versa.  Mimicking sampling from the general population
    arrange(x)
#arrange because helps to visualize the data, but not an essential step - arrange doesn't actually change your results
    
}
```

Check if this works...

```{r}
boot_sample(sim_df_nonconst) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm") +
  ylim(-5, 16)
#fitted line moves slightly with each bootstrap sample

#next, can I actually estimate that intercept and slope, and save those across bootstrap samples??
```

```{r}
boot_sample(sim_df_nonconst) %>% 
  lm(y ~ x, data = .) %>% 
  broom::tidy()
#next, can I actually estimate that intercept and slope, and save those across bootstrap samples??

```

